Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-08-29 paper/paper.tex
text editing in paper Tested everything on quipu.

\section{Comparison of Mannik's Experimental and Our Stadium Shapes}
We have made and studied plots and movies of the MinD in a number of
simulated cell shapes and sizes. We have found that the deterministic
simulations of the protein reaction sequence discussed above results
in robust oscillatory behavoir of polar selection and oscillation in
not only symmetrical shapes, such as the wild type pill shape, but
also in very assymetrical, flatttened cells, such as those studied by
Mannik.  This differs with Mannik's experimental findings that show
much less regular behavoir, with a large variation in the location of
maxima.  On the other hand, our stochastic simulations show both
irregular behavoir in the assymetrical shapes and more regular, pole
to pole oscillations in the symmetric pill and stadium shaped cells,
agreeing well with Mannik's experimental findings.

Figure~\ref{arrow-pill-stadium} and Figure~\ref{randst-plot-ave} show color
plots of time-averaged MinD overlayed by arrows depicting sequential
maxima that are global in space and local in time.  Our shapes that
are designed to look like Mannik cells show well the inadequacy of the
deterministic method of simulation and conversely the accuracy of the
stochastic method.  Our stochastic method movies (and arrow plots of
Figure~\ref{randst-plot-ave}) show irregular location of maxima that is
similar to what is observed experimentally by Mannik. There is a
general bias towards back and forth behavoir, which can be seen in the
box plots of Figure~\ref{box-mannik}, but the behavoir is irregular and
the location of the maxima vary a significant amount.  On the other
hand, the deterministic simulations show behavoir that is much too
regular when compared with experiment, exhibiting locations of maxima
that do not vary.  The deterministic method is seen to be inadequate.

@@

Our simulations of pill shapes show behavoir that is once again well
matched by experiment, but this time the deterministic method seems to
be as aduaquet as the stochastic in giving accurate results.  The
experimental cells show regular back and forth behavoir, as do both
simulations.  In this case the natural bias of the deterministic
method towards robustness and regularity works in its favor.

@@

We find that both the stochastic and the deterministic solutions show
regular back and forth oscillations the stadium shaped cells, as can
be seen Figure~\ref{box-stadium}.  This suggests that it is the irregular,
assymetric shape of the Mannik cells, not their size or their
flattened nature, that causes these cells to exhibit irregularity in
MinD oscillations.

@@

Figure~\ref{randst-plot-ave} shows a comparison of irregular shapes
simulated with the stochastic method and with the deterministic
method, along with a plot taken from Mannik showing the location of
sequential MinD maxima in the experimental cells.  Like Mannik's, the
tips of the arrows in our plots show subsequent maxima, and the data
overlays a color plot of time averaged MinD concentration.  These
plots illustrate the adequacy of the stochastic method and inadequacy
of the deterministic one.  The stochastic plots show similar random
behavoir to Manniks

@@

\section{Interpretation of Data}
\section{Conclusion}
\bibliography{paper}

@@

%% \subsubsection{Randst-97-}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-600-9700-1500}
%%   %%\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9600-1500}
%%   \caption{Total protein fluctuation in the sideways flying saucer
%%     pancake cell shape.  The vertical axis shows stacked the total
%%     number of proteins that are of four different compound stages and
%%     in the different sections of the cell.}
%%   \label{box-97}
%% \end{figure}

%% Early on in its simulation the sideways flying saucer pancake shape
%% shows a clear breakdown of an unstable equilibrium oscillation.  We
%% have set the intial conditions so that MinD density is highly
%% concentrated to the left of a dividing line that runs parrallel to the
%% left vertical side.  For the first number of oscillations the maxima
%% trade off right to left, from a high peak in the rightward middle
%% compartment to vertically opposed peaks in the poles of the long axis
%% of the cell.  This is seen in the first part of Figure~\ref{box-97},
%% where the lower and upper sections peak in unison, and the middle
%% right section peaks $\pi$ radians removed from these.  After a few
%% oscillations the back and forth oscillations tip toward the vertical
%% direction and the system loses its initial oscillation pattern.  It
%% falls quickly instead into a back and forth pattern between the two
%% opposing endcaps.  The oscillation at this point is reminiscent of the
%% pills shape oscillation, once again showing the robust nature of the
%% back and forth, two pole pattern.

%% During each period, directly after one of the polar maxima has
%% subsided, there is a small spike with the regular maxima pattern (a
%% spike in membrane bound MinD:ATP immediately followed by a spike in
%% membrane bound MinD:MinE:ATP) seen in the small middle right
%% compartment section.  The animations show that the MinD, oscillating
%% back and forth between the upper and lower poles, seems to catch and
%% lag in the indentation for a bit before moving on.  Once again the
%% positive radius of curvature acts as an attractive place for MinD to
%% cluster on the walls, and it is not released to move on through the
%% cytoplasm again until MinE has formed its ring and completely reacted
%% it away.

%% \subsubsection{Randst-96-Star}
%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-700-700-9600-1500}
%%   %%\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9600-1500}
%%   \caption{Total protein fluctuation in the star cell shape.  The vertical axis shows stacked the total
%%     number of proteins that are of four different compound stages and
%%     in the different sections of the cell.}
%%   \label{box-96}
%% \end{figure}

%% This clover shape finds a regular oscillatory behavoir after about
%% seven periods that consists of two maxima that peak regularly in the
%% corners accross from each other at the same times (Figure
%% \ref{box-96}) and then again opposing maxima that peak at the same
%% time in the other two corners.\fixme{should try some different
%%   starting conditions with this to see if there's another patter.}

%% Looking at at few periods in detail.  We find that there is a small
%% increase in cytoplasmic MinE always directly before the beggining
%% of climb in the membrane bound MinE peak.  Also, there is at any one
%% time a very large amount more of minE that is membrane bound then
%% cytoplasmic in any ones section.  A difference of a factor of roughly
%% 20 when membrane MinE is at its peak, and roughly 12 when membrane
%% MinE is at it's minimum.  There seem to be two modes of oscillation,
%% each exhibiting a density peak maxima that moves back and forth
%% diagonaly from one corner to the opposite, as if there were two
%% overlapping, criss-crossed cells oscillating.  In the box plot this is
%% seen as roughly coinciding membrane minD peaks in the lower left and
%% right corners, followed one pi later in peaks in the upper right and
%% upper left corners.  Interesting that this did not exhibit
%% counterclockwise motion that Huangs 2008 paper might suggest.  Should
%% run sims that start in the wacky way that just started running with
%% triangles.
%% \newline

%% We haven't simulated long enough yet, so no real pattern has shown
%% itself.  There's certainly oscillations, and it's almost as if there's
%% competing modes - top to bottom, left to right, corner to corner, etc.
%% Interesting.

%% \fixme{maybe?}The animations seem to relax into a pattern in which the oscillations
%% go from the smallest corner, out to the other three, and then back to
%% the smallest corner?  \fixme{There isn't enough data to see properly
%%   though, need to run for longer.}

%% The double pole oscillation pattern, in which one pole is located in a
%% single corner of the triangle and the other pole is spread out between
%% the opposite two, can be seen in Figure \ref{box-triangle}.  The
%% equilateral triangle shows a spike in both the Center and Left
%% sections at the same time, both showing the familiar sequence of peaks
%% amoungst the different states of the protein seen in the poles of the
%% pill shaped bacteria.  A spike in the other corner, showing the common
%% sequence, occurs during the minima of the former two, wth a higher
%% concentration at the peak.  A similar pattern is seen in the
%% iscosolese triangular shape, where interestingly the onset of the
%% maxima in the single maxima corner (the Left section) builds more
%% quickly than it fades.  This is presumably becuase as the cytoplasmic
%% minD:ATP diffuses down a more accute corner, it is sourrounded by
%% closer cell walls than it would be in a less acute corner, so that the
%% it is more apt at any diffusive step to make contact and cluster on
%% the walls.  That particular part of the process is therefore
%% accelerated, driving the build up of the maxima forward.

%% The right triangle shows a similar pattern except that here there is
%% an assymetry between the two corners that maximize at roughly the same
%% time.  One corner is further from the single maximizing corner and it
%% consisitently peaks later.  The MinD simply have further to travel
%% before they maximize there.

%% For these three sizess and shapes there is not sign that the
%% oscillation patterns once they are set will break down.

%% \begin{figure}
%% %%  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-600-600-600-1500}
%%   \caption{Total protein fluctuation of an equilateral triangle with
%%     larger dimensions (each side is roughly $6\micron$.  The vertical
%%     axis shows stacked the total number of proteins that are of four
%%     different compound stages and in the different sections of the
%%     cell. The time limits have been set to show the broken down
%%     section of the simulation.}
%%   \label{box-triangle}
%% \end{figure}

%% The larger equilateral shape shows the same pattern as the smaller
%% version does only until fifteen periods, at which point there is a
%% breakdown of the oscillation pattern.  During the last three stable
%% periods, the single maximizing corner, the lower left section in the
%% figure, has a lower peak each period as the membrane bound MinD:ATP in
%% the other corners lag more and more before being eaten away by the
%% MinE.  Finally these maxima lag so much that they break down the
%% regular pattern.  There then occurs about five periods of semi chaotic
%% behavoir, until it settles back into another stable oscillation that
%% is similar to the original one, except that a different corner has
%% been chosen to be the singlular maxima (the right section in the
%% figure), while the other two corners maximize at the same time.  This
%% interesting breakdown and recovery shows both the robustness of the
%% double pole oscilatory tendency and the importance of required
%% distance of travel for the different proteins.

>>>>

\section{Comparison of Mannik's Experimental and Our Stadium Shapes}
\fixme{should I say anything about different/large sizes?}
We have made and studied plots and movies of the MinD in a number of
simulated cell shapes and sizes. We have found that the deterministic
simulations of the protein reaction sequence discussed above results
in robust oscillatory behavoir of polar selection and oscillation in
not only symmetrical shapes, such as the wild type pill shape, but
also in very assymetrical, flatttened cells, such as those studied by
Mannik.  This differs with Mannik's experimental findings that show
much less regular behavoir, with a large variation in the location of
maxima.  On the other hand, our stochastic simulations show both
robust pole to pole oscillations in the symmetric cells and irregular
behavoir in the assymetrical shapes, agreeing well with Mannik's
experimental findings.

Figure~\ref{arrow-pill-stadium} and Figure~\ref{randst-plot-ave} show
color plots of time-averaged MinD overlayed by arrows depicting
sequential maxima that are global in space and local in time.  Our
stochastic method movies (and arrow plots of
Figure~\ref{randst-plot-ave}) show irregular location of maxima that
is similar to what is observed experimentally by Mannik. There is a
general bias towards back and forth behavoir, which can be seen in the
box plots of Figure~\ref{box-mannik}, but the behavoir is irregular
and the location of the maxima vary a significant amount.  On the
other hand, the deterministic simulations show behavoir that is much
too regular when compared with experiment, exhibiting locations of
maxima that do not vary.  The deterministic method is seen to be
inadequate.

@@

@@

We find that both the stochastic and the deterministic solutions show
regular back and forth oscillations in the stadium shaped cells, as
can be seen Figure~\ref{box-stadium}.  This suggests that it is the
irregular, assymetric shape of the Mannik cells, not their size or
their flattened nature, that causes these cells to exhibit
irregularity in MinD oscillations.

@@

@@

\section{Interpretation of Data}
\section{Conclusion}
No conclusion :(
\bibliography{paper}

@@
